---
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir ="docs") })
title: "Tutorial Document"
author: "Yida Wu"
date: "`r format(Sys.Date())`"
CJKmainfont: SimSun
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: vignette
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup,include = FALSE}
knitr::opts_chunk$set(
  fig.align = 'center',
  fig.show="asis", 
  fig.align="center", 
  out.width = "80%",
  message = FALSE,
  warning = FALSE)
```

This tutorial document is used to reproduce the results of **Real data analysis**.

# R package
Load required R packages.

```{r}
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(purrr)
library(snowfall)
library(Matrix)
library(hash)
library(igraph)
library(ggraph)
library(RColorBrewer)
library(ggtext)
library(lemon)
library(ggpubr)
library(caret)
library(cowplot)
```

# Set parameters
```{r}
allsubtree = c("AB","C","D","E","MS")
maxdelta_time = 5
```

# Set path
```{r}
hufilepath = "./data/184onset/"
oldfilepath = "./data/184rawdata/"
expdatapath = "./data/expdata/"
figurepath = "./figure/"
readypath = "./data/readydata/"
realresultmainpath = "./real_result/"
if(!dir.exists(realresultmainpath)){
  dir.create(realresultmainpath)
  for(sub_dir in allsubtree){
    output_dir=file.path(realresultmainpath,sub_dir)
    if (!dir.exists(output_dir)){
    dir.create(output_dir)
    }
  }
}
simuresultmainpath = "./simu_result/"
if(!dir.exists(simuresultmainpath)){
  dir.create(simuresultmainpath)
  for(sub_dir in allsubtree){
    output_dir=file.path(simuresultmainpath,sub_dir)
    if (!dir.exists(output_dir)){
    dir.create(output_dir)
    }
  }
}
```

# R function
```{r}
source("./code/dataIntegrationFun.R") # for data integration
source("./code/preScreenFun.R") # for pre-screening process
source("./code/mainMCMCFun.R") # for MCMC
source("./code/pictureFun.R") # for Network visualization 
```

# Data preparation and integration
## Data preparation
Choose the gene expression onset cells with their descendants in each raw data file.
```{r}
allfilename = read_csv(file="data/filenameinfo.csv")
oldfilename = allfilename$oldfilename
hufilename = allfilename$hufilename
newfilename = allfilename$newfilename
sapply(1:184,express_data,oldfilename,hufilename,newfilename,hufilepath,oldfilepath,expdatapath)
```

## Data integration
After running the below code, we acquire five binary pairwise time series subtree datasets which are stored in the `readypath`. For a comprehensive overview of these five subtree datasets, please refer to Table 1 in the paper.
```{r}
for(numk in 1:5){
  subtree = allsubtree[numk]
  realpath = paste(readypath,subtree,"/",sep="")
  allchoosedata = cellnumStat(subtree,expdatapath,realpath)
  allfilename = read_csv(file="data/filenameinfo.csv")
  usedata = read_csv(file=paste(realpath,"all_",subtree,"_cell.csv",sep=""),show_col_types = FALSE)
  geneCopyFun(subtree,usedata,allfilename,realpath)
  load(file=paste(realpath,subtree,"_simu_info.RData",sep=""))
  genecopy_h = simu_info$genecopy
  genename = simu_info$genename
  copyname = simu_info$copyname
  cellname = simu_info$cellname
  Mc = simu_info$Mc
  interData(cellname,copyname,Mc,expdatapath,subtree,maxdelta_time,realpath)
  load(paste(realpath,subtree,"_allfun.RData",sep=""))
  realdata = read.csv(file = paste(realpath,subtree,"_",maxdelta_time,"_condata.csv",sep=""),
                      check.names=FALSE)
  realdata = as_tibble(realdata)
  hashData(allfun,realdata,copyname,cellname,maxdelta_time,Mc,realpath,subtree)
  load(paste(realpath,subtree,"_",maxdelta_time,"_allhash.RData",sep=""))
  computeM(allhash,subtree,maxdelta_time,realpath)
  load(paste(realpath,subtree,"_M_",maxdelta_time,"_.RData",sep=""))
  GeneDataset = realdata[,-c(1:6)]
  InfoDataset = realdata[,c(1:6)]
  mergeCopyFun(allhash,genecopy_h,M,realdata,GeneDataset,InfoDataset,subtree,deletecellrate,deletegenerate,copyname,cellname,mergerate,realpath)
}
```

# BBTD for real datasets
## Pre-screening process

Apply pre-screening process to screen the $a_{ij}$.

```{r}
p=1 # set seed
theta = 0.05
for(numk in 1:5){
  subtree = allsubtree[numk]
  realpath = paste(readypath,subtree,"/",sep="")
  load(file=paste(realpath,subtree,"_simu_info.RData",sep=""))
  Mc = simu_info$Mc
  eachdelta = 1/(round(Mc)-1)
  maxdelta = maxdelta_time*eachdelta
  newdata = read_csv(file=paste(realpath,"mer_",subtree,"_",maxdelta_time,"_selectdata.csv",sep=""),show_col_types = FALSE)
  finalinfo = newdata[,1:6]
  finaldata = newdata[,-c(1:6)]
  finaldata = map_dfc(finaldata,function(x) replace_na(x,0))
  selectgene = names(finaldata)
  load(paste(realpath,"mer_",subtree,"_",maxdelta_time,"_allchooserow.RData",sep=""))
  load(paste(realpath,"mer_",subtree,"_",maxdelta_time,"_origenehash.RData",sep=""))
  aijindex = list()
  for(i in 1:length(selectgene)){
    aijindex[[i]] = setdiff(1:length(selectgene),i)
  }
  names(aijindex) = selectgene
  realaijmat = 1
  diffData(finaldata,finalinfo,allchooserow,subtree,p,realpath)
  lastvalueFun(p,selectgene,allchooserow,finalinfo,maxdelta,eachdelta,newallhash,realpath)
  preScreen(aijindex,selectgene,finaldata,finalinfo,allchooserow,realaijmat,
                              maxdelta_time,Mc,theta,ifallpvalue=T,ifreal=T,realpath,p,subtree,newallhash)
}
```

## MCMC

After performing the pre-screening process, MCMC algorithm is implemented with five
independent MCMC chains (5 cpu cores) for each subtree.

```{r}
p=1 # set seed
subtree = "D" 
#"C","D","E","MS"
mcmcnum = 5
num_maxiters = 8
iters = 4
betasd = 0.1
alphasd = 0.1
lambdasd = 0.1
kasd = 0.01
realpath = paste(readypath,subtree,"/",sep="")
realresultpath = paste(realresultmainpath,subtree,"/",sep="")
realMCMC(subtree,p,mcmcnum,num_maxiters,iters,betasd,alphasd,lambdasd,kasd,maxdelta_time,realpath,realresultpath)
```

## Network visualization
After running the below code, we can get Figure 5 in the paper, which is saved in `figurepath`.
```{r}
multiNetFun(allsubtree,p,maxdelta_time,readypath,realresultmainpath,figurepath)
```
